Evolutionary demography and the maintenance of sexually antagonistic polymorphism in outcrossers and simultaneous hermaphrodites

### 2020-11-23 doc/SA-Hermaphrodites-wDemography.tex
Minor edits & new dataDeltaPolySpace data
Minor cleanup over weekend.

Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single bi-allelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites. The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing. A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

The population state at time $t$ is described by a stage $\times$ genotype distribution matrix, which can be transformed into a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by genotype, then stage, then by how individuals were produced (self vs.~outcrossing). For a single locus with two alleles ($A$ and $a$), we have genotypes $g \in \{AA,\, Aa,\, aa\}$, giving the population state vector:
\begin{linenomath*}
\begin{equation}
	\tilde{\mbf{n}}(t) =  \left[
								\begin{array}{c}
									\mbf{n}^{S}_{AA}(t) \\
									\mbf{n}^{S}_{Aa}(t) \\
									\mbf{n}^{S}_{aa}(t) \\ \hline
									\mbf{n}^{X}_{AA}(t) \\
									\mbf{n}^{X}_{Aa}(t) \\
									\mbf{n}^{X}_{aa}(t) \\ 
						\end{array} \right],
\end{equation}
\end{linenomath*}

\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ (where $i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively. The proportional population vector is given by
\begin{linenomath*}
\begin{equation} \label{eq:propPopVec}
	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|} = \mbf{p}(t),
\end{equation}
\end{linenomath*}

@@

The fertility matrices through male and female function are
\begin{linenomath*}
\begin{equation}\label{eq:FS}
	\mbf{F}_{i} = \left(
					\begin{array}{cc}
						0 & f_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}
\noindent and
\begin{linenomath*}
\begin{equation}\label{eq:FX}
	\mbf{F}^{\prime}_{i} = \left(
					\begin{array}{cc}
						0 & f^{\prime}_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}

@@

\begin{table}[htbp]
 \centering
 \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
\label{tab:Fitness}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
Female function:	& $f_{AA}$		& $f_{Aa} - h_f s_f$	& $f_{aa} - s_f$ \\	
Male function:		& $f^{\prime}_{AA} - s_m$& $f^{\prime}_{Aa} - h_m s_m$	& $f^{\prime}_{aa}$ 		\\	
\hline
\end{tabular}
\end{table}

>>>>

Here, we briefly derive a matrix model incorporating multiple life-cycle stages and a single diallelic locus under SA selection on male and female sex functions for a population of partially selfing simultaneous hermaphrodites. The derivation and some of the key results follow closely those presented in \citet{deVriesCaswell2019a} and \citet{deVriesCaswell2019b}; and in fact, the model presented here reduces to the two-sex stage-structured model of \citet{deVriesCaswell2019b} under obligate outcrossing. A full derivation of the model and analyses is presented in \hl{Appendix A}, and all simulation code necessary to reproduce the results are available at \url{https://github.com/colin-olito/SA-Hermaphrodites-wDemography}.

@@

The population state at time $t$ is described by a stage $\times$ genotype distribution matrix, which can be transformed into a population state vector, $\tilde{\mbf{n}}(t)$, which is ordered by genotype, then stage, then by how individuals were produced (self vs.~outcrossing). For a single diallelic locus with alleles $A$ and $a$, we have three genotypes ($AA,\, Aa,\, aa$; $g = 3$), giving the population state vector:
\begin{linenomath*}
\begin{equation}
	\tilde{\mbf{n}}(t) =  \left[
								\begin{array}{c}
									\mbf{n}^{S}_{AA}(t) \\
									\mbf{n}^{S}_{Aa}(t) \\
									\mbf{n}^{S}_{aa}(t) \\ \hline
									\mbf{n}^{X}_{AA}(t) \\
									\mbf{n}^{X}_{Aa}(t) \\
									\mbf{n}^{X}_{aa}(t) \\ 
						\end{array} \right],
\end{equation}
\end{linenomath*}

\noindent where $\mbf{n}^{S}_{i}$ and $\mbf{n}^{X}_{i}$ ($i \in \{AA,Aa,aa\}$) are the stage $\times$ genotype distribution vectors of individuals produced by self-fertilization and outcrossing, respectively. The proportional population vector is given by
\begin{linenomath*}
\begin{equation} \label{eq:propPopVec}
	\tilde{\mbf{p}}(t) = \frac{\tilde{\mbf{n}}(t)}{ \| \tilde{\mbf{n}}(t) \|} = \mbf{p}(t),
\end{equation}
\end{linenomath*}

@@

The fertility matrices through female and male function are
\begin{linenomath*}
\begin{equation}\label{eq:FS}
	\mbf{F}_{i} = \left(
					\begin{array}{cc}
						0 & f_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}
\noindent and
\begin{linenomath*}
\begin{equation}\label{eq:FX}
	\mbf{F}^{\prime}_{i} = \left(
					\begin{array}{cc}
						0 & f^{\prime}_{i} \\
						0 & 0
					\end{array}
				\right),
\end{equation}
\end{linenomath*}

@@

\begin{table}[htbp]
 \centering
 \caption{\bf Fitness expressions for Sexually Antagonistic selection in hermaphrodites}
\label{tab:Fitness}
\begin{tabular}{lccc}
 \toprule
					&  \multicolumn{3}{c}{{\textit{Genotype}}} \\ 
\cline{2-4}
					& $AA$			& $Aa$ 					& $aa$ 		\\ \hline
Female function:	& $f_{AA}$		& $f_{Aa}(1 - h_f s_f)$	& $f_{aa}(1 - s_f)$ \\	
Male function:		& $f^{\prime}_{AA}(1 - s_m)$& $f^{\prime}_{Aa}(1 - h_m s_m)$	& $f^{\prime}_{aa}$ 		\\	
\hline
\end{tabular}
\end{table}
